---
title: "Module 4 Topic 11 Example"
output:
  html_document: 
    toc: TRUE
  pdf_document: default
---


Read in the data
```{r}
prostate <- read.csv('prostatecancer.csv') # reads in .csv file

# convert binary variables into factors
prostate$xray <- factor(prostate$xray)
prostate$size <- factor(prostate$size)
prostate$grade <- factor(prostate$grade)
prostate$ni <- factor(prostate$ni)
```

## Fitting Logistic Regression Model

```{r}
summary(fit1 <- glm(ni ~ grade, family = binomial(link = "logit"), data = prostate))
```

## Estimating the OR from logit link

Recall in the Topic 10 example where we obtained the coefficient OR using `coef()` within the `exp()` function in order to undo the logit transformation on the coefficient estimates.

```{r}
exp(coef(fit1)) # Odds Ratio of coefficients
exp(fit1$coefficients) # 2nd method
```

To obtain the 95% confidence interval was also introduced in the Topic 10 example using `confint.default()`
```{r}
exp(confint.default(fit1))
```


We can also use the custom function from Topic 10 if you wish to obtain the usual summary output but w/ OR.

```{r, include = FALSE}
logit_summary <- function(x){
  stopifnot(class(x)[1] == "glm") # input must be of class 'glm'
  
  preds <- unlist(strsplit(as.character(x$formula[3]), # extract predictors used
                           split = "[[:space:]]\\+[[:space:]]"))
  LL <- stats::logLik(x) # log likelihood
  y <- as.character(x$formula[2]) # outcome variable
  tStat <- with(x, null.deviance - deviance) # chi-square test statistic
  AIC <- x$aic # Akaike information criterion
  BIC <- (-2 * LL) + (log(length(x$residuals)) * (length(preds) + 1)) # Bayesian Information Criterion
  pval <- with(x, stats::pchisq(null.deviance - deviance, # p-value of model
                         df.null - df.residual, lower.tail = FALSE))
  
  mod_stats <- merge(summary(x)$coefficients, exp(confint.default(x)), by = "row.names") # model stats
  mod_stats$`Odds Ratio` <- exp(mod_stats$Estimate) # add 'Odds Ratio'
  mod_stats <- subset(mod_stats, select = -Estimate) # drop 'Estimate'
  mod_stats <- mod_stats[,c(1, ncol(mod_stats),2:(ncol(mod_stats)-1))] # reorder columns
  

  
  
  tbl <- data.frame(nrow = length(preds), ncol = 5) # data.frame
  
  output <- list(LL, y, tStat, AIC, BIC, pval, tbl) # list of diagnostics
  names(output) <- c("log likelihood", "outcome", "LR chi2",
                     "AIC", "BIC", "Prob > chi2","results") # names for list
  
  output$results <- mod_stats
  
  return(output)
}
```

```{r}
logit_summary(fit1)
```


## Logistic Regression

For base R there is only one way to perform a logistic regression, there are no alternatives. We only use the `glm()` function. More information about `glm()` can be found using the `help(glm)` and more on logistic regression can be found [here](https://stats.idre.ucla.edu/r/dae/logit-regression/).

### Covariate Adjustment

```{r}
logit_summary(fit2 <- glm(ni ~ grade + size, family = binomial(link = "logit"), data = prostate))
```

To perform a covariate adjustment comparing different grades with different sizes we will be using the `glht()` function from the `multcomp` package to fit a new model.

```{r}
summary(fit3 <- multcomp::glht(fit2, linfct = c("grade1 - size1 = 0")))
```

Once again keep in mind that the provided `Estimate` is transformed and you would need to apply the `exp()` function to obtain the estimated OR of `r exp(coef(fit3))`, $\exp(-0.6959) = 0.4986$.

The same applies for the 95% confidence interval.
```{r}
exp(confint.default(fit3))
```

Information on how to test contrasts in R using the `multcomp` package can be found [here](https://stats.idre.ucla.edu/r/faq/how-can-i-test-contrasts-in-r/) and documentation on how to use the `glht()` function can be found using `help(glht, multcomp)`. I highly recommend examining these so you can know how to apply this towards the Data Analysis 4 if necessary.

### Predicted Probability

```{r}
exp(-1.6008+0.7293)/(1+exp(-1.6008+0.7293))
```

To use R for calculating the estimated probability
```{r}
prostate$prob <- predict(fit2, type = 'response')
subset(prostate, size == 0 & grade == 1, select = c(grade, size, prob))
```

Currently we are unable to mimic the `STATA` margins command on slide 13 to get specific predicted probabilities. Instead we will use the `predict.glm()` function, which requires us to enter new data of the same form as the data used to fit the model.

```{r}
newData <- data.frame(size = '0', grade = '1', stringsAsFactors = T)
predict.glm(fit2, newdata = newData, type = 'response')
```


### Interpreting OR with a Continuous Predictor

This is as simple as just fitting the model using a continuous predictor
```{r}
summary(fit4 <- glm(ni ~ acid, family = binomial(link = "logit"), data = prostate))
logit_summary(fit4)
```

### Estimating OR with Other Than 1 Unit Of Change

If we want to use a change on `acid` predictor of 0.5 then the following is applicable
```{r}
summary(fit5 <- multcomp::glht(fit4, linfct = c("0.5 * acid = 0")))
cbind(exp(coef(fit5)), exp(confint.default(fit5))) # OR and 95% confidence interval
```