---
title: "Module 2 Topic 6 Example"
output:
  html_document: default
---


Read in the Module 6 Example 1 Data

```{r}
mod2_ex1 <- read.csv("Module2Topic6Ex1.csv")
```

Use the `str` function to get a description of the data
```{r}
str(mod2_ex1)
```

Calculate the quadratic term `p2`
```{r}
mod2_ex1$p2 <- mod2_ex1$percent^2
summary(mod2_ex1$percent) # to obtain a summary of the `percent` variable
```

Now to center both the `percent` and `p2` variables
```{r}
mod2_ex1$pcent <- mod2_ex1$percent - mean(mod2_ex1$percent)
mod2_ex1$pcent2 <- mod2_ex1$pcent^2
```

Look at correlation matrix
```{r}
cor(mod2_ex1)
```

To graph a scatterplot matrix in R we can use the `plot` function
```{r}
plot(mod2_ex1)
```

There is also the `GGally` package which can make a more visually striking scatterplot matrix
```{r}
library(GGally)
ggpairs(mod2_ex1)
```


To calculate VIF will require the `car` package
```{r}
fit1 <- lm(csat ~ percent + p2 + high, data = mod2_ex1)
summary(fit1)

library(car)
vif(fit1)
```

To mimic the STATA output for 1/VIF and Mean VIF, we will write a simple convenience function.

```{r}
f_calculate_vif <- function(fit) {
  v <- c(v <- car::vif(fit))
  m <- cbind(v, 1/v)
  colnames(m) <- c("VIF", "1/VIF")
  print(m)
  cat("Mean VIF: ", mean(v))
}

f_calculate_vif(fit1)
```

```{r}
fit2 <- lm(csat ~ pcent + pcent2 + high, data = mod2_ex1)
summary(fit2);
f_calculate_vif(fit2)
```

Now for the Body Fat Percentage Data

```{r}
mod2_ex2 <- read.csv("Module2Topic6-bodyfat.csv")
fit3 <- lm(bodyfat ~ midarm + thigh + triceps, data = mod2_ex2)
summary(fit3)
f_calculate_vif(fit3)
```

```{r}
cor(mod2_ex2)
```

```{r}
fit4 <- lm(bodyfat ~ triceps, data = mod2_ex2)
summary(fit4)
```

```{r}
fit5 <- lm(bodyfat ~ thigh, data = mod2_ex2)
summary(fit5)
```

```{r}
fit6 <- lm(bodyfat ~ midarm + thigh, data = mod2_ex2)
summary(fit6); 
f_calculate_vif(fit6)
```

There are numerous packages in CRAN available for easy identification of other diagnostic collinearity statistics. One of them would be the `mctest` package

```{r}
library(mctest)
imcdiag(fit3)
```

