---
title: "Module 2 Topic 5 Example"
output:
  html_document: default
---


Read in the same data from Module 1

```{r}
data_mod1 = read.csv("Module1Topics1and2Example.csv")
```


Fit the model as before with the previous example
```{r}
fit <- lm(y ~ x1 + x2 + x3, data = data_mod1)
summary(fit)
```

```{r}
data_mod1$yhat <- predict(fit) # predict yhat
```

```{r}
library(MASS)
data_mod1$jack <- studres(fit) # jack residuals
```

```{r}
data_mod1$lev <- hatvalues(fit) # Calculate leverage
```

```{r}
data_mod1$dist <- cooks.distance(fit) # Calculate Cook's D
```

```{r}
data_mod1$dfit <- dffits(fit) # Calculate DFFITS
```

```{r, eval = FALSE}
dfbeta(fit) # Calculates DFBETAS
```

```{r}
# DFBETAs for each covariate
data_mod1$dx1 <- dfbeta(fit)[, 'x1']
data_mod1$dx2 <- dfbeta(fit)[, 'x2']
data_mod1$dx3 <- dfbeta(fit)[, 'x3']
```

```{r}
data_mod1$resid <- resid(fit) # Calculate OLS Residuals
```

Our complete dataframe
```{r}
head(data_mod1)
```

Now that we have calculated the above diagnostic statistics we can then move on to creating graphs using the `ggplot2` package.

```{r}
library(ggplot2)
p <- ggplot(data_mod1)

p + aes(x = lev) +
  geom_boxplot(colour = 'darkblue', fill = 'cornflowerblue') +
  ggtitle("Boxplot of Leverage Values") +
  xlab("Leverage") +
  coord_flip() +
  theme_bw()
```

```{r}
p + aes(x = patient, y = lev) +
  geom_point(colour = 'darkblue') +
  ggtitle('patient Plot of Leverage Values') +
  theme_bw()
```

```{r}
p + aes(x = dist) +
  geom_boxplot(colour = 'darkblue', fill = 'cornflowerblue') +
  ggtitle("Boxplot of Cook's D") + 
  xlab("Cook's Distance") +
  coord_flip() +
  theme_bw()
```

```{r}
p + aes(x = patient, y = dist) +
  geom_point(colour = 'darkblue') +
  ggtitle("Patient Plot of Cook's D") +
  theme_bw()
```

```{r}
p + aes(x = dfit) +
  geom_boxplot(colour = 'darkblue', fill = 'cornflowerblue') +
  ggtitle("Boxplot of DFFITS") + 
  xlab("DFFITS") +
  coord_flip() +
  theme_bw()
```

```{r}
p + aes(x = patient, y = dfit) +
  geom_point(colour = 'darkblue') +
  ggtitle("Patient Plot of DFFITS") +
  theme_bw()
```

Below is a boxplot of the DFBETAS for the `x1` variable

```{r}
p + aes(x = dx1) +
  geom_boxplot(colour = 'darkblue', fill = 'cornflowerblue') +
  coord_flip() +
  ggtitle('Boxplot of x1 DFBETAS') +
  theme_bw()
```

To create a single boxplot of all of the covariate DFBETAS for comparison we must reshape our dataframe appropriately using the `tidyr` package
```{r}
library(tidyr)
melted_df <- pivot_longer(data_mod1, cols = c('dx1', 'dx2', 'dx3'), names_to = "variable") 
head(melted_df)
```

Now we can plot as usual
```{r}
p2 <- ggplot(melted_df)

p2 + aes(x = value, fill = variable) +
  geom_boxplot() +
  xlab('DFBETA') +
  ggtitle("Boxplot of DFBETAS") +
  coord_flip() +
  theme_bw()
```

It is very difficult to view the range of `dx3` so it would be wise to make a separate boxplot for `dx3` just like w/ `x1` above.


```{r}
p + aes(x = resid^2, y = lev) +
  geom_point(colour = 'darkblue') +
  ggtitle("Leverage Residual Plot") +
  xlab('Normalized Residual Squared') +
  theme_bw()
```


```{r}
p + aes(x = yhat, y = jack) +
  geom_point(shape = 1, aes(stroke = 1.1, size = dist), colour = 'darkblue') +
  ggtitle("Jack Knife Residuals vs Fits with Circles Proportional to Cooks D") +
  xlab("Fitted Values") +
  ylab("Studentized Residuals") +
  geom_hline(yintercept = 0, colour = 'red') + 
  theme(legend.title = element_blank()) +
  theme_bw()
  
```

In order to sort our observations by Cook's Distance we can rearrange our dataframe
```{r}
sorted_df <- data_mod1[order(data_mod1$dist), ] # create new data.frame sorted by 'dist' variable
head(sorted_df)
```

To identify the points greater than the cutoff value
```{r}
(cut_off <- 4 / nrow(data_mod1)) # calculate Cook's D cutoff using 4/n formulation
high_dist <- sorted_df[sorted_df$dist >= cut_off, ] # identifies patient values equal or greater than cut_off
high_dist
```

As we can see, it was the 6th and 44th observations that had a Cook's D greater than `r cut_off` which was our cutoff.


To refit the model while removing observations with whichever you determined to be the cutoff point for Cook's D
```{r}
data_mod1 <- data_mod1[!data_mod1$patient %in% high_dist$patient, ] # new data.frame w/o high Cook's D
fit2 <- lm(y ~ x1 + x2 + x3, data = data_mod1)
summary(fit2)
```

Now to make predictions again using the new model
```{r}
data_mod1$yhat2 <- predict(fit2)
```

Now lets compare the yhat values between the models
```{r}
p3 <- ggplot(data_mod1)

p3 + aes(x = yhat, y = yhat2) +
  geom_point(colour = 'darkblue') +
  ggtitle("Comparison of Two Fits w/ Influential Points Removed") +
  theme_bw()
```

To find the correlation
```{r}
cor(data_mod1$yhat, data_mod1$yhat2)
```

To perform a robust regression requires using the `MASS` package
```{r}
# library(MASS)
MASS::rlm(formula = y ~ x1 + x2 + x3, data = data_mod1)
```

**IMPORTANT**

There will be a difference in the coefficient values between the `rlm()` function of R and `rreg` of STATA, despite that both functions are performing similar statistical procedures, both are using just one of many different forms of robust regression. Nonetheless, for this class these differences are negligible.