---
title: "Module 1 Topic 1 Example"
output:
  html_document: default
---


We can read in Stata data files in directly with the `read_stata()` function from `library(haven)`. Alternatively, we can read **C**omma **S**eparated **V**alues files with `read.csv()`

```{r}
data_mod1 = read.csv("Module1Topics1and2Example.csv")
```

List the data
```{r}
data_mod1
```

We will regress y on x1 and x2
```{r}
fit = lm(y ~ x1 + x2, data = data_mod1)
summary(fit)
```

Compute fitted values. You can look at the arguments for `predict.lm()` with `?predict.lm`
```{r}
predict(fit)
```
Add the fitted to our data.frame
```{r}
data_mod1$yhat = predict(fit)
```

List the fitted values for patient == 5
```{r}
data_mod1[data_mod1$patient == 5,]
```

We can predict the values when x1 = 2.8, x2 = 7. With base R, you will have to calculate the 

```{r}
predict(fit, newdata = data.frame(x1 = 2.8, x2 = 7), interval = "confidence", se = TRUE)
```

When x1 = 2.8, x2 = 5.
```{r}
predict(fit, newdata = data.frame(x1 = 2.8, x2 = 5), interval = "confidence", se = TRUE)
```
The $R^2$ for models 1, 2, and 3 are printed by the `summary.lm()` method.

```{r}
summary(lm(y ~ x1, data = data_mod1))
```

```{r}
summary(lm(y ~ x2, data = data_mod1))
```


```{r}
summary(lm(y ~ x1 + x2, data = data_mod1))
```


We can center variables in R manually

```{r}
data_mod1$c_x1 = data_mod1$x1 - mean(data_mod1$x1)
data_mod1$c_x2 = data_mod1$x2 - mean(data_mod1$x2)
```

Alternatively, we can use the `scale()` function to center the variables `x1` and `x2` 

```{r}
x1x2_centered = scale(data_mod1[,c("x1", "x2")], center = TRUE, scale = FALSE)
head(x1x2_centered)
```

We can see that the two methods produce the same values.

```{r}
data_mod1$c_x1 == x1x2_centered[,1]
data_mod1$c_x2 == x1x2_centered[,2]
```

Perform the regression on the full model using the centered variables.

```{r}
summary(fit2 <- lm(y ~ c_x1 + c_x2, data = data_mod1))
```


Now generate the predicted values for the centered regression.


```{r}
data_mod1$c_yhat = predict(fit2)
```

Compare the predicted values

```{r}
data_mod1[,c("yhat", "c_yhat")]
```

