---
title: "Module 3 Topic 8 Example"
output:
  html_document: default
---


Read in the data and fit the max model and check collinearity diagnostics
```{r}
df <-  read.csv("Module3Topic8ExData.csv")

summary(fit <- lm(y ~ ., data = df))

# Re-use the vif function from topic 6
f_calculate_vif <- function(fit) {v <- c(v <- car::vif(fit));m <- cbind(v, 1/v);colnames(m) <- c("VIF", "1/VIF");print(m);cat("Mean VIF: ", mean(v))}
f_calculate_vif(fit)
```

Show the correlation matrix
```{r}
cor(df)
```
There are numerous model selection libraries available to R. Within base R, there are stepwise selection is implemented in `step()`.

Exhaustive search, forward or backward stepwise, or sequential replacement can be performed with `library(leaps)`. Other popular model selection libraries include `bestglm`, `glmulti`.

`library(HH)` contains some helper functions that makes the outputs from `library(leaps)` easier to digest.

```{r}
library(leaps)
library(HH)
summaryHH(fit <- regsubsets(y ~ . , data = df, method = "exhaustive"))
```
```{r}
# Choose model 2
summary(lm.regsubsets(fit, 2))
```
Backward and forward selection can be performed with the `regsubsets()` function, but we will demonstrate stepwise selection with base R.

Backward selection by AIC (selection by p-value is supported by `leaps::regsubsets`)
```{r}
step(lm(y ~ ., data = df), scope = list(upper = ~., lower = ~1), direction = "backward")
```
Forward selection by AIC
```{r}
# Foward
step(lm(y ~ 1, data = df), scope = list(upper = ~., lower = ~1), direction = "forward")
```
Chunkwise selection is not supported by `step()`. A quick Comprehensive R Archive Network (CRAN) seems to indicate that stepwise/chunkwise model selection may not be implemented in any user implemented packages.

Backwards Elimination Locking in the first term (x6)
```{r}
step(lm(y ~ ., data = df), scope = list(upper = ~., lower = ~x6), direction = "backward")
```

```{r, message=FALSE}
library(ggplot2)
fit = lm(y ~ x1 + x3, data = df)
qplot(predict(fit), MASS::studres(fit),
      main = "Jackknife Residuals vs Fitted Values",
      xlab = "Fitted Values", ylab = "Jackknife Residuals") +
  theme_bw() +
  geom_hline(aes(yintercept = 0), color = "red")
```

Assess homogeneity of variance assumption.

An implementation of the Breusch-Pagan Test can be found in `library(car)` as `ncvTest()`. 

```{r, message=FALSE}
library(car)
ncvTest(fit)
```

Plot the kernel density
```{r}
plot(density(studres(fit)), ylim = c(0, 0.4), col = "blue")
lines(seq(-3, 3, 0.01), dnorm(seq(-3, 3, 0.01)), col = "red")
```

Test for non-normality with Shapiro-Wilks
```{r}
shapiro.test(studres(fit))
```
