---
title: "DA7_2021_R_Key"
author: "Arvon Clemons II"
date: "4/23/2021"
output:
  pdf_document:
    toc: yes
  html_document:
    df_print: paged
    toc: yes
---

## Libraries
```{r}
library(ggplot2)
library(MASS)
```

## Import Data
```{r}
DA6 <- read.csv("./da6.csv")

# factor categorical variables
DA6$health <- factor(DA6$health)
DA6$sex <- factor(DA6$sex)
DA6$adldiff <- factor(DA6$adldiff)
DA6$race <- factor(DA6$race)
DA6$privins <- factor(DA6$privins)
```

```{r Custom Functions, include = FALSE}
log_summary <- function(x, robust = FALSE, irr = FALSE){
  stopifnot("glm" %in% class(x), # input must be of class 'glm'
            is.logical(robust)) # robust argument must be logical
  
  waldChi2 <- function(x){
    # Get the coefficients of the non-intercept parameters 
    b <- matrix(coef(x)[2:length(coef(x))])
    # Get the covariance matrix of terms in b
    V <- vcov(x)[2:nrow(vcov(x)),2:nrow(vcov(x))]
    # Create the R matrix for the 3 linear hypothesis
    R <- matrix(0, nrow = nrow(b), ncol = nrow(b))
    # Only the diagonal in R needs to be filled in
    diag(R) <- 1
    # Create a vector of values Rb = r
    r <- 0
    # Calculate Wald statistics
    chi2 <- as.numeric(t((R%*%b - r)) %*% (solve(R %*% V %*% t(R))) %*% (R%*%b - r))
    
    return(chi2)
  }
  
  preds <- unlist(strsplit(as.character(x$formula[3]), # extract predictors used
                           split = "[[:space:]]\\+[[:space:]]"))
  LL <- stats::logLik(x) # log likelihood
  y <- as.character(x$formula[2]) # outcome variable
  tStat <- if(robust == FALSE){
    with(x, null.deviance - deviance) # LR chi-square test statistic
  }
  else(
    waldChi2(x) # Wald chi-square test statistic
  )
  McF <- signif(1 - stats::logLik(x)/stats::logLik(glm(as.formula(paste(y, "1", sep = "~")), # McFadden's Pseudo R^2
                                                       family = poisson(link = "log"), data = x$data)), digits = 4)
  AIC <- stats::AIC(x) # Akaike information criterion
  BIC <- stats::BIC(x) # Bayesian Information Criterion
  pval <- round(with(x, stats::pchisq(tStat, # p-value of model
                                      df.null - df.residual, lower.tail = FALSE)), digits = 4)
  Disp_P <- sum(residuals(x, type = "pearson")^2) / x$df.residual # Pearson Dispersion
  
  mod_stats <- merge(summary(x)$coefficients, confint.default(x),
                     by = "row.names", sort = FALSE) # model stats
  
  if(robust == TRUE){
    rse <- sqrt(diag(sandwich::vcovHC(x, type="HC0"))) # Robust Standard Error
    
    mod_stats <- subset(mod_stats,
                        select = -3) # drop 'Std. Error'
    
    # robust statistics
    mod_stats$`z value` <- round(coef(x) / rse, digits = 2)
    mod_stats$`Pr(>|z|)` <-  round(2 * stats::pnorm(abs(stats::coef(x)/rse), lower.tail=FALSE), 
                                   digits = 3)
    mod_stats$`2.5 %` <- stats::coef(x) - stats::qnorm(0.975, lower.tail = T) * rse
    mod_stats$`97.5 %` <- stats::coef(x) + stats::qnorm(0.975, lower.tail = T) * rse
    mod_stats$`Robust Std. Error` <- rse # add 'Robust Std. Error'
    mod_stats <- mod_stats[,c(1:2, ncol(mod_stats),3:(ncol(mod_stats)-1))] # reorder columns
  }
  
  if(irr == TRUE){
    mod_stats$Estimate <- exp(mod_stats$Estimate)
    mod_stats$`2.5 %` <- exp(mod_stats$`2.5 %`)
    mod_stats$`97.5 %` <- exp(mod_stats$`97.5 %`)
    # mod_stats$`Robust Std. Error` <- msm::deltamethod(list(~ exp(x1), ~ exp(x2), ~ exp(x3), ~ exp(x4)), coef(x),
    #                                sandwich::vcovHC(x, type="HC0"))
    
    names(mod_stats) <- c("Row.names", "IRR", "Delta Std. Error",
                          "z value", "Pr(>|z|)", "2.5 %", "97.5 %")
  }
  
  tbl <- data.frame(nrow = length(preds), ncol = 5) # data.frame
  
  output <- list(LL, y, tStat, McF, AIC, BIC, pval, Disp_P, tbl) # list of diagnostics
  names(output) <- c("log likelihood", "outcome", "chi2", "Pseudo R^2",
                     "AIC", "BIC", "Prob > chi2","(1/df) Pearson","results") # names for list
  
  output$results <- mod_stats
  
  return(output)
}

# To Do
## Fix Robust Std. Error using Delta Method when arg 'irr = TRUE'
## Fix Pseudo R^2 when using Poisson Regression model w/ offset

lincom <- function(x, linfct, exp = FALSE, ...){
  stopifnot("glm" %in% class(x), # input must be of class 'glm'
            is.logical(exp))  # robust argument must be logical
  
  linComb <- multcomp::glht(x, linfct, ...)
  linSummary <- summary(linComb)
  df <- setNames(data.frame(Estimate = linSummary$test$coefficients,
             sigma = linSummary$test$sigma,
             z = linSummary$test$tstat,
             pval= linSummary$test$pvalues),
           c("Estimate", "Std. Err", "z", "P>|z|"))
  Confint <- stats::confint(linSummary)$confint[,2:3]
  Confint <- matrix(Confint, ncol = 2)
  colnames(Confint) <- c("2.5 %", "97.5 %")
  df <- cbind(df, Confint)
  
  if(exp == TRUE){
    df <- apply(df[c(1, 5:6)], 2, exp)
  }
  return(df)
}

poisson_GOF <- function(x, type = c("deviance","pearson")){
  if(missing(type)){
    type <- "deviance" # default arg 'type = "deviance"'
  }
  stopifnot(class(x)[1] == "glm", # input must be of class 'glm'
            class(type) == "character", # type arg must be 'character'
            type == "deviance" | type == "pearson")
  
  if(type == "deviance"){
    # Deviance based likelihood ratio test: $\chi^2_D$ 
    cat("Goodness-of-fit chi2 = ", x$deviance,
        paste0("\nProb > chi2(",x$df.residual,") = "), pchisq(x$deviance, df=x$df.residual, lower.tail = F))
  }
  if(type == "pearson"){
    # Pearson goodness of fit: $\chi^2_P$
    y <- as.character(x$formula[2]) # outcome variable
    y_j <- x$y
    e_ep_j <- predict(x, type = "response")
    cat("Goodness-of-fit chi2 = ", sum((y_j - e_ep_j)^2/e_ep_j),
        paste0("\nProb > chi2(",x$df.residual,") = "),
        pchisq(sum((y_j - e_ep_j)^2/e_ep_j), df=x$df.residual, lower.tail = F))
  }
}

lrt <- function(reduced, full){ # important to provide reduced model in correct argument
  calc <- stats::anova(reduced, full)
  LR_chi <- calc$Deviance[2] # Test Statistic
  DF <- calc$Df[2] # degree of freedom
  
  pval <- round(stats::pchisq(q = LR_chi, df = DF, lower.tail = F), digits = 4)
  
  cat(paste0("chi2(", DF,")=",signif(LR_chi, digits = 4),
      "\nProb > chi2 = ", pval))
  
  # result <- list(LR_chi, DF, pval)
  # names(result) <- c("LR chi2", "df", "Prob > chi2")
  # return(result)
}
```

Answer the following questions and justify your answers.  Interpretations are necessary!

## Question 1:

Using glm with family poisson and log link, fit a Poisson model that estimates the expected number of physician office visits adjusting for health status, gender, race, condition of limiting activities of living, race, private insurance information, age, chronic conditions, and education. Assess the scale parameters and determine if the assumptions of the Poisson model you fit are violated. In addition, assess if you think there are excess zeros. Explain your answer.

```{r}
fitPOI <- glm(visit ~ health + sex + adldiff + race + privins + age + cond + edu,
            family = poisson(link = "log"), data = DA6)

log_summary(fitPOI)
```

```{r}
dispersion_parameter <- sum(residuals(fitPOI,type ="pearson")^2)/fitPOI$df.residual
dispersion_parameter
```


The estimated Pearson chi-square dispersion parameter is `r log_summary(fitPOI)$"(1/df) Pearson"` >> 1, indicating the presence of overdispersion so a negative binomial model would be more appropriate.

They might also fit a negative binomial regression to assess this which is ok

```{r}
fit1 <- glm.nb(visit ~ health + sex + adldiff + race + privins + age + cond + edu, data = DA6)
summary(fit1)
```

```{r}
# Likelihood Ratio Test
pchisq(2 * (logLik(fit1) - logLik(fitPOI)), df = 1, lower.tail = F)
```

This indicates the Negative Binomial Regression model is more appropriate than the Poisson Regression model.

To assess the excess zeros,

```{r}
sum(DA6$visit == 0) / nrow(DA6)
```

The percentage of zero count is about 5.6%. Zero-inflated models do not need to be considered.

## Question 2

Using negative binomial regression refit the model, perform a likelihood ratio test or Wald test to identify whether health status, gender, race, condition of limiting activities of living, race, private insurance information, age, chronic conditions, and education are univariably associated with the mean number of physician office visits. (This means to fit single models, one for each of those covariates).

```{r}
fit2 <- update(fit1, . ~ health)
fit3 <- update(fit1, . ~ sex)
fit4 <- update(fit1, . ~ race)
fit5 <- update(fit1, . ~ adldiff)
fit6 <- update(fit1, . ~ privins)
fit7 <- update(fit1, . ~ age)
fit8 <- update(fit1, . ~ cond)
fit9 <- update(fit1, . ~ edu)
fitNull <- update(fit1, . ~ 1)
```

```{r}
FitList <- mget(ls(pattern = "fit[[:digit:]]")) # list of fits 1 - 9
lapply(FitList, anova, fitNull)[-1]
```

**So we will not include sex and age.**

## Question 3

Fit a multivariable model that describes the data given the information from #2 (don’t consider interactions) and complete the following with appropriate interpretation:


```{r}
fit10 <- update(fit1, . ~ . - age - sex)
summary(fit10)
```

Adjusted for the other independent variables, what is the estimated ratio of 
number of physician office visits for 50 year old males who reported an excellent health condition and 2 chronic conditions to 50 year old females who reported an average health condition and 5 chronic conditions.

```{r}
lincom(fit10, linfct = "health3 - health2 + (-3 * cond) = 0", exp = T)
```

## Question 4

Perform an appropriate test or other analysis to assess the appropriateness of the negative binomial model in #3, comment on those results.

Scale parameters from the model in #3 are around 1, so negative binomial is ok to use.  Could also refit using nbreg to the get LRT of alpha=0,

```{r}
# dispersion parameter
sum(residuals(fit10, type = "pearson") ^ 2) / fit10$df.residual
```

Dispersion parameter `r sum(residuals(fit10, type = "pearson") ^ 2) / fit10$df.residual` is close to 1 so NB is ok. Or using LRT

```{r}
# Likelihood Ratio Test
pchisq(2 * (logLik(fit10) - logLik(update(fitPOI, . ~ . - age - sex))), df = 1, lower.tail = F)
```

Test of overdispersion using the LRT of alpha = 0 is significant, indicating the NB model is more appropriate.

## Question 5

Compute the adjusted residuals, predicted visits and influence statistics from the model in #3.  Generate the graph of the adjusted residuals vs the fitted visits weighted by the influence using:

```{r, eval= FALSE}
p <- ggplot(DA6)

p + aes(x = vhat, y = adjres, size = cooksd) +
  geom_point(colour = "darkblue") +
  geom_hline(yintercept = 0, colour = "darkred") +
  labs(x = "Predicted mean visit") + 
  theme_bw()
```

```{r}
DA6$resid <- resid(fit10, type = "pearson")
DA6$hat <- hatvalues(fit10)
DA6$adjres <- with(DA6, resid / sqrt(1 - hat))
DA6$cooksd <- cooks.distance(fit10)
DA6$vhat <- predict(fit10, type = "response")
DA6$index <- 1:nrow(DA6)
DA6 <- DA6[, c(ncol(DA6), 2:(ncol(DA6) - 1))]
```

```{r}
p <- ggplot(DA6)

p + aes(x = vhat, y = adjres, size = cooksd) +
  geom_point(colour = "darkblue") +
  geom_hline(yintercept = 0, colour = "darkred") +
  labs(x = "Predicted mean visit") + 
  theme_bw()
```

We see the fanning shape which we would expect for these data.  However, there does seem to be some very large residuals as well as influential points.